### 2014-10-29 contract-language.tex
Spacing edits to motivation, contract-language. More text in transactions.

\item \textbf{Eventually consistency}: Eventually consistent operations can
  be satisfied as long as the client can reach at least one replica. For
  example, \cf{deposit} is an eventually consistent operation; its semantics
  does not require its action to manifest on all replicas before other
  operations in its session are allowed to proceed. While eventually
  consistent data store typically offer \emph{basic} eventual consistency
  with all possible anomalies, we assume that our store provides stronger
  semantics that remain highly-available~\cite{BailisHAT,COPS}; the store
  always exposes a causal cut of the updates. This semantics can be formally
  captured in terms of the following contract definition:
  \vspace{-0.8em}
  \begin{smathpar}
  \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}
  \noindent where $\small \hboZ = ((\soZ \cap \sameobjZ) \cup \visZ)^+$.

\item \textbf{Causal consistency}: Causally consistent operations
  require to see a causally consistent snapshot of the object state,
  including the actions performed on the same session.  The latter
  requirement entails that if two operations $o_1$ and $o_2$ from the
  same session are applied to two different replicas $r_1$ and $r_2$,
  the second operation cannot be discharged until the effect of $o_1$ is
  merged with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance}
  operation requires causal consistency, as it requires the operations
  from the same session to be visible, which cannot be guaranteed under
  eventual consistency. We assume that causality is only tracked through
  operations on the same object; two operations in the same session but
  on different objects are considered causally unrelated under this
  definition. Stores typically avoid tracking causality across objects
  to mitigate overheads when causality tracking is unnecessary. The
  corresponding store semantics is captured by the contract $\ccc$
  defined below:
  \vspace{-0.8em}
  \begin{smathpar}
  \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

\item \textbf{Strong Consistency}: Strongly consistent operations may block
  indefinitely under network partitions. An example is the total-order
  contract on \cf{withdraw} operation. The corresponding store semantics is
  captured by the $\scc$ contract definition:
  \vspace{-0.8em}
  \begin{smathpar}
  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
  \end{smathpar}

@@

Our goal is to map application-level consistency constraints of
operations to appropriate store-level consistency guarantees capable
of satisfying constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:
\begin{definition}
A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
\begin{center}
\end{center}
\end{definition}

\vspace{-2em} \noindent The $\Delta$ referred in the above defintion captures
assumptions about the nature of primitive relations, such as $\Rvis$
and $\Rso$ are irreflexive, and the happens-before relation $\hbZ$ is
acyclic, preventing thin-air reads.

>>>>

\item \textbf{Eventually consistency}: Eventually consistent operations can
  be satisfied as long as the client can reach at least one replica. For
  example, \cf{deposit} is an eventually consistent operation; its semantics
  does not require its action to manifest on all replicas before other
  operations in its session are allowed to proceed. While eventually
  consistent data store typically offer \emph{basic} eventual consistency
  with all possible anomalies, we assume that our store provides stronger
  semantics that remain highly-available~\cite{BailisHAT,COPS}; the store
  always exposes a causal cut of the updates. This semantics can be formally
  captured in terms of the following contract definition:

  \vspace{-1em}
  \begin{smathpar}
  \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}
  \noindent where $\small \hboZ = ((\soZ \cap \sameobjZ) \cup \visZ)^+$.

\item \textbf{Causal consistency}: Causally consistent operations
  require to see a causally consistent snapshot of the object state,
  including the actions performed on the same session.  The latter
  requirement entails that if two operations $o_1$ and $o_2$ from the
  same session are applied to two different replicas $r_1$ and $r_2$,
  the second operation cannot be discharged until the effect of $o_1$ is
  merged with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance}
  operation requires causal consistency, as it requires the operations
  from the same session to be visible, which cannot be guaranteed under
  eventual consistency. We assume that causality is only tracked through
  operations on the same object; two operations in the same session but
  on different objects are considered causally unrelated under this
  definition. Stores typically avoid tracking causality across objects
  to mitigate overheads when causality tracking is unnecessary. The
  corresponding store semantics is captured by the contract $\ccc$
  defined below:

  \vspace{-1em}
  \begin{smathpar}
  \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

\item \textbf{Strong Consistency}: Strongly consistent operations may block
  indefinitely under network partitions. An example is the total-order
  contract on \cf{withdraw} operation. The corresponding store semantics is
  captured by the $\scc$ contract definition:

  \vspace{-1em}
  \begin{smathpar}
  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
  \end{smathpar}

@@

Our goal is to map application-level consistency constraints of
operations to appropriate store-level consistency guarantees capable
of satisfying constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:

\begin{definition}
A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
\begin{center}
\end{center}
\end{definition}

\vspace{-1em}
\noindent The $\Delta$ referred in the above defintion captures
assumptions about the nature of primitive relations, such as $\Rvis$
and $\Rso$ are irreflexive, and the happens-before relation $\hbZ$ is
acyclic, preventing thin-air reads.

### 2014-10-29 transactions.tex
Spacing edits to motivation, contract-language. More text in transactions.

While making both transactions serializable would ensure correctness,
distributed stores rarely offer serializable transactions. Moreover,
serializability is unavailable and hinders scalability. As we will see, these
transactions can be satisfied with much weaker isolation guarantees. Despite
the atomicity offered by the transaction, anomalies are still possible.

>>>>

While making both transactions serializable would ensure correctness,
distributed stores rarely offer serializable transactions. Moreover,
serializability is unavailable and hinders scalability. As we will see, these
transactions can be satisfied with much weaker isolation guarantees. Despite
the atomicity offered by the transaction, anomalies are still possible. For
example, the two \cf{getBalance} operations in \cf{totalBalance} transactions
might be served by different replicas with distinct set of committed \cf{save}
transactions. If the first(second) \cf{getBalance} operation witness a
\cf{save} transaction that is not witnessed by the second(first)
\cf{getBalance} operation, then the balance returned will be less(greater) than
the actual balance. It is not immediately apparent which weakest isolation
guaratee will be sufficient to prevent the anomaly.

Instead, \name requires the programmer to simply state the consistency
requirement as a contract. Since we would like both the \cf{getBalance}
operations to witness the same set of \cf{save} transactions, we define
$\psi_2$ as:

\begin{smathpar}
\begin{array}{l}
\cv_{2} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, (c:\rcf{withdraw} \vee \rcf{deposit}), d. \\
\qquad \txn{a,b}{c,d} ~\wedge~ \vis{d}{b} ~\wedge~ \sameobj{a}{c} \Rightarrow \vis{c}{a}
\end{array}
\end{smathpar}
